---
title: "individual1"
author: "Molly Buckler"
date: "11/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
```

```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

head(acc)
head(per)
```

##Questions

##1) Are there some days of the week where more accidents happen than the others (see FARS manual, use variable DAY_WEEK)?
```{r}
week <- table(acc$DAY_WEEK)
week
```

##There are some days of the week were more accidents happen than the others. Most of the accidents are concentrated on friday, saterday and sunday.

##2) Create a data frame containing the persons who are fatally hurt in the accidents (look up variable INJ_SEV)
```{r}
fate <- per$INJ_SEV
head(fate)
```

##3)Create a data frame containing the most dangerous vehicle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make. Make sure to handle the missing values appropriately. (look up variable MAKE)
```{r}

```

##4)Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the most dangerous vehicle, and what you find from the map. (Hint: Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication. Use readxl::read_xlsx to read in the GLCs.)
```{r}
```

##5) Join the accident and person table (work out which variable(s) to use)
```{r}
total <- merge(acc, per,by=c("COUNTY", "DAY", "MONTH", "HOUR", "MINUTE"))
head(total)

```

##6) Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results.
```{r}
#whs <- total %>% select(total$DAY_WEEK, total$HOUR, total$SEX)
#head(whs)
whs <- total[ , c("DAY_WEEK", "HOUR", "SEX")]
head(whs)
#whsCount <- count(whs, c("DAY_WEEK", "HOUR", "SEX"))
whsTally <- whs %>% add_tally()
head(whsTally)

```

##7) Now plot a choropleth map of the number of deaths on a county level.
```{r}
#install.packages("readxl")
#library("readxl")
#USA <- read_excel("USA.xlsx")
#head(USA)

```

##8) Is summer or winter more dangerous? Does this depend on states? Explore and explain.
```{r}
danger <- total[ , c("MONTH", "INJ_SEV")]
head(danger)
nd <- aggregate(INJ_SEV ~ MONTH, data = total, sum)
head(nd)
```
